---
title: "Parent Exclusion"
author: "R. Mark Sharp"
date: "8/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(parentfindr)
library(rmsutilityr)

```

## Input of Allele File and Trios File

Define allele and trios file names.

```{r define-allele-and-trio-filenames}
alleleFile <- "../inst/testdata/snp-animal-alleles.xlsx"
triosFile <- "../inst/testdata/snp-potential-trios.xlsx"
#alleleFile <- "inst/testdata/snp-animal-alleles.txt"
#triosFile <- "inst/testdata/snp-potential-trios.txt"
#alleleFile <- "inst/testdata/str-animal-alleles.txt"
#triosFile <- "inst/testdata/str-potential-trios.txt"
#alleleFile <- "inst/testdata/str-animal-alleles.xlsx"
#triosFile <- "inst/testdata/str-potential-trios.txt"
#alleleFile <- "../inst/testdata/str-animal-alleles_rms.xlsx"
#triosFile <- "../inst/testdata/str-potential-trios.txt"

```

## Data Format and Questions

I have been reading through the code and have a question regarding how missing 
loci (alleles) and invalid loci (alleles) are calculated.
Only the left of the left and right pairs of the locus geneotypes are used for 
scoring, which makes me assume that if the left one is missing, 
both are missing. Is that correct?
A parent’s locus is counted as __missing__ if it does not have a value in the 
left allele even if this locus is missing for the offspring.
A parent’s locus is counted as __invalid__ if it does have a value in the left 
allele but the offspring does not have a value in the left allele.
This allows the calculation of the denominator for the percent non-discrepant 
alleles.

## Hard Coded Test Data

```{r set-up-test-data}
  minNumber <- 4
  maxDiscrepant <- 4
  thMissing <- 4
  thDiscrepant <- 4
  microsatellites <- TRUE
  siresFirst <- FALSE
  selDiscrepant <- TRUE
  nAlleles <- 0
  dateType = "YYYYMMDD" # is one of c("YYYYMMDD", "mm/dd/YYYY")
  firstParentType <- "dams" # is one of c("dames", "sires")

```



```{r read-in-example-data, echo=TRUE}
trios <- getTrios(triosFile)
animalAlleles <- getAnimalAlleles(alleleFile, dateType)

```

```{r calculate-scores}
scores <- getScores(trios, animalAlleles)
scoresDf <- getScoresDf(scores, firstParentType)
scoresDf <- scoresDf[]
```
